---
title: "CVD and Diabetes by Race"
output: html_document
date: '2022-12-29'
---
# Install Packages

```{r loadPackages, message = FALSE}
library(tidyverse)
library(readr)
```

# Load Data

```{r loadData, message = FALSE}
hispanic_df <- read_csv("Data/raceData/Hispanic.csv")
white_df <- read_csv("Data/raceData/Non_hispanic_white.csv")
black_df <- read_csv("Data/raceData/Non_hispanic_black.csv")
asian_df <- read_csv("Data/raceData/Asians.csv")
```

# Tidy Data 
 Column "Race" is added because there is not a column or row for race in the
 original CSVs, as each CSV only included data from one race.
 Column "...1" is renamed for clarity.

```{r tidyData, message = FALSE}
clean_asian_df <- mutate(asian_df, Race = "Asian") %>%
  rename(Question = ...1) %>%
  filter(Question %in% c(
    "Diagnosed with any CVD",
    "Ever diagnosed with diabetes W5",
    "HIGH SENSTVTY C-RCTV PROTN (hsCRP)(MG/L)"
    ))

clean_black_df <- mutate(black_df, Race = "Black") %>%
  rename(Question = ...1) %>%
  filter(Question %in% c(
    "Diagnosed with any CVD",
    "Ever diagnosed with diabetes W5",
    "HIGH SENSTVTY C-RCTV PROTN (hsCRP)(MG/L)"
    ))

clean_hispanic_df <- mutate(hispanic_df, Race = "Hispanic") %>%
  rename(Question = ...1) %>%
  filter(Question %in% c(
    "Diagnosed with any CVD",
    "Ever diagnosed with diabetes W5",
    "HIGH SENSTVTY C-RCTV PROTN (hsCRP)(MG/L)"
    ))

clean_white_df <- mutate(white_df, Race = "White") %>%
  rename(Question = ...1) %>%
  filter(Question %in% c(
    "Diagnosed with any CVD",
    "Ever diagnosed with diabetes W5",
    "HIGH SENSTVTY C-RCTV PROTN (hsCRP)(MG/L)"
    ))
```

# Join Dataframes 
```{r joinDataframes, message = FALSE}
asian_and_black_df <- rbind(clean_asian_df, clean_black_df)
hispanic_and_white_df <- rbind(clean_hispanic_df, clean_white_df)
all_races_df <- rbind(asian_and_black_df, hispanic_and_white_df)
```

# Visualize Boxplot 

## Omitted Data Values 
Whenever limiting the range for the "HIGH SENSTVTY C-RCTV PROTN (hsCRP)(MG/L)" variable, we decided to omit 
the maximum values since, the data states that at most, participants of the survey had hsCRP levels between 75 and 205 (mg/L), which is highly unlikely, as they generally result from extremely severe and rare cases of inflammation. 
Therefore, we deicded to keep the range of the boxplot between 0 and 10, since this is where 75% of the data
fell between. 

```{r crpByRace, message = FALSE}
all_races_df_CRP <- all_races_df %>% 
  slice(3,6,9,12)

# Note: The matrix must be race x summary statistics in order of (min, Q1, Q2, Q3, max)
#         [min] [q1] [q2] [q3] [max]
# [Asian]
# [Hispanic]
# [Black]
# [White]


raceStats <- list(stats=t(cbind(all_races_df_CRP$Minimum,
                 all_races_df_CRP$`Percentile 25`,
                 all_races_df_CRP$Median,
                 all_races_df_CRP$`Percentile 75`,
                 all_races_df_CRP$Maximum)),
               n=c(all_races_df_CRP$Count[1],
                   all_races_df_CRP$Count[2],
                   all_races_df_CRP$Count[3],
                   all_races_df_CRP$Count[4]))
  
bxp(raceStats, show.names = FALSE, main="Measure of hsCRP by Race (male and female)", xlab="hsCRP(mg/L)", ylab="Race", ylim = c(0,10), horizontal = TRUE)
axis(2, at = 1:4, labels = c("Asian", "Hisp.", "Black", "White"))
```

# Visualize 
```{r visualizeDataframe, message = FALSE}
# https://www.statology.org/ggplot2-barplot-with-multiple-variables/

all_races_df_bar <- all_races_df %>% 
  slice(1,2,4,5,7,8,10,11)

ggplot(all_races_df_bar, aes(fill=Question, y=Mean, x=Race)) +
  geom_bar(position='dodge', stat='identity') +
  ggtitle('CVD & Diabetes Diagnosis by Race') +
  labs(x = "Race", y = "Percentage of Sample Diagnosed") +
  scale_fill_manual('Diagnosis', values=c('firebrick3','cornflowerblue'))
  
```