---
title: "CRP vs CVD and Diabetes Incidence"
author: "Randy Truong"
date: "2022-12-25"
output: html_document
---
# Installing Packages 

```{r loadPackages, message=FALSE}
library(tidyverse)
library(readr)
```

# Data Setup  

## TODOS: 
-> Import all of the CSVs - **DONE**   
-> Reduce the DFs to essential rows - **DONE**   
  - CVD Diagnosis    
  - Diabetes diagnosis    
  - log(hsCRP)   
  - Abnormally high CRP rates    
  - # of Observations      
        
-> Create a race column - **DONE**    
-> Join all of the datasets together - **DONE**   
-> Create Plots     

## Important Terminology: 
CVD -> Cardiovascular Disease           
CRP -> C-Reactive Protein           

```{r loadData, message=FALSE}
mAsianDF <- read_csv("./Data/mData/M_Asian.csv")
mHispanicDF <- read_csv("./Data/mData/M_hispanic.csv")
mBlackDF <- read_csv("./Data/mData/M_non-hispanic_black.csv")
mWhiteDF <- read_csv("./Data/mData/M_non-hispanic_white.csv")
mDF <- read_csv("./Data/mData/allMales.csv")
```

# Filtering of Data Sets 

In the datasets, we mostly just care about the rows "Diagnosed with CVD" (1),  
"Ever diagnosed with diabetes W5" (11), "log10(hsCRP) W4" (24), "High hsCRP (>3 mg/L) W4" (25), and "Observations" (34).

```{r filterMaleAsian, message=FALSE}

# Remember to rename first row 
mAsianDFClean <- mutate(mAsianDF, Race = "Asian") %>% 
  slice(1, 11, 23, 24, 25, 34) 

mHispanicDFClean <- mutate(mHispanicDF, Race = "Hispanic") %>% 
  slice(1, 11, 23, 24, 25, 34) 

mBlackDFClean <- mutate(mBlackDF, Race = "Black") %>% 
  slice(1, 11, 23, 24, 25, 34) 

mWhiteDFClean <- mutate(mWhiteDF, Race = "White") %>% 
  slice(1, 11, 23, 24, 25, 34) 

mDFClean <- mDF %>% 
  slice(1, 11, 23, 24, 25, 34)

```

```{r createGiantFrame, message=FALSE}
raceList1 <- rbind(mAsianDFClean, mHispanicDFClean)
raceList2 <- rbind(mBlackDFClean, mWhiteDFClean)

finalRaceListM <- rbind(raceList1, raceList2)

```


# Data Visualization  

## TODOS: 

-> Visualize hsCRP rate by race     
-> Visualize the frequency of CVD and the frequency of diabetes diagnoses    
-> Compare log(hsCRP) to CVD / Compare log(hsCRP) to Diabetes    

```{r generic}
# data
d1 <- c(1,2,3,4,5)
d2 <- c(1,2,3,5,8,13,21,34)

# summaries (generated with quantile and structured accordingly)
z1 <- list(
    stats=matrix(quantile(d1, c(0.05,0.25,0.5,0.75,0.85))),
    n=length(d1)
)
z2 <- list(
    stats=matrix(quantile(d2, c(0.05,0.25,0.5,0.75,0.85))),
    n=length(d2)
)

# merging the summaries appropriately
z <- list(
    stats=cbind(z1$stats,z2$stats),
    n=c(z1$n,z2$n)
)

# check result
print(z)

# call bxp with needed parameters ("at" can/should also be used here)
bxp(z=z,varwidth=TRUE)
```

```{r crpByRace}
finalRaceListM_CRP <- finalRaceListM %>% 
  slice(3, 9, 15, 21)

mStats <- list(stats=t(cbind(finalRaceListM_CRP$Minimum,
                      finalRaceListM_CRP$`Percentile 25`,
                      finalRaceListM_CRP$Median,
                      finalRaceListM_CRP$`Percentile 75`,
                      finalRaceListM_CRP$Maximum)),
          n=c(finalRaceListM_CRP$Count[1],
              finalRaceListM_CRP$Count[2],
              finalRaceListM_CRP$Count[3],
              finalRaceListM_CRP$Count[4]))

print(mStats)
bxp(mStats, show.names = FALSE, main="Measure of hsCRP by Race (male)", ylab= "Race", xlab="hsCRP(mg/L)", ylim= c(0,10), horizontal = TRUE )
axis(2, at = 1:4, labels = c("Asian", "Hispanic", "Black", "White")) # '2' means left axis 
# axis(4) # '4' means bottom axis 

```

```{r diseaseFrequency}
# Frequency of Cardiovascular Disease (CVD) diagnoses 
# Bar graph 

```

```{r crpByDisease,echo=FALSE}
# Relationship between hsCRP and Cardiovascular Disease
# Bar graph 

# Relationship between hsCRP and Diabetes
# Bar graph 

```


