---
title: "(F) CRP vs CVD and Diabetes Incidence"
author: "Randy Truong"
date: "2022-12-29"
output: html_document
---

# Installing Packages 

```{r loadPackages, message=FALSE}
library(tidyverse)
library(readr)
```

# Data Setup  

## TODOS: 
-> Import all of the CSVs - **DONE**   
-> Reduce the DFs to essential rows - **DONE**   
  - CVD Diagnosis    
  - Diabetes diagnosis    
  - log(hsCRP)   
  - Abnormally high CRP rates    
  - # of Observations      
        
-> Create a race column - **DONE**    
-> Join all of the datasets together - **DONE**   
-> Create Plots     

## Important Terminology: 
CVD -> Cardiovascular Disease           
CRP -> C-Reactive Protein           

```{r loadData, message=FALSE}
fAsianDF <- read_csv("./Data/fData/F_Asian.csv")
fHispanicDF <- read_csv("./Data/fData/F_hispanic.csv")
fBlackDF <- read_csv("./Data/fData/F_non-hispanic_black.csv")
fWhiteDF <- read_csv("./Data/fData/F_non-hispanic_white.csv")

```

# Filtering of Data Sets 

In the datasets, we mostly just care about the rows "Diagnosed with CVD" (1),  
"Ever diagnosed with diabetes W5" (11), "log10(hsCRP) W4" (24), "High hsCRP (>3 mg/L) W4" (25), and "Observations" (34).

```{r filterMaleAsian, message=FALSE}

# Remember to rename first row 
fAsianDFClean <- mutate(fAsianDF, Race = "Asian") %>% 
  slice(1, 11, 24, 25, 34) 

fHispanicDFClean <- mutate(fHispanicDF, Race = "Hispanic") %>% 
  slice(1, 11, 24, 25, 34) 

fBlackDFClean <- mutate(fBlackDF, Race = "Black") %>% 
  slice(1, 11, 24, 25, 34) 

fWhiteDFClean <- mutate(fWhiteDF, Race = "White") %>% 
  slice(1, 11, 24, 25, 34) 

```

```{r createGiantFrame, message=FALSE}
raceList1 <- rbind(fAsianDFClean, fHispanicDFClean)
raceList2 <- rbind(fBlackDFClean, fWhiteDFClean)

finalRaceList <- rbind(raceList1, raceList2)

```


# Data Visualization  

## TODOS: 

-> Visualize hsCRP rate by race     
-> Visualize the frequency of CVD and the frequency of diabetes diagnoses    
-> Compare log(hsCRP) to CVD / Compare log(hsCRP) to Diabetes    

```{r crpByRace}
# Frequency of hsCRP between different races
# Box plot 

# Current Issue: Trying to visualize the data given the five-number summary
# Resolution: Use bxp? 

# https://www.rdocumentation.org/packages/graphics/versions/3.6.2/topics/bxp

finalRaceList_CRP <- finalRaceList %>% 
  slice(3, 8, 13, 18) 

CRP_Graph <- ggplot(finalRaceList_CRP, aes(x = Race, y = Mean)) + 
  geom_boxplot()

CRP_Graph

```