---
title: "(F) CRP vs CVD and Diabetes Incidence"
author: "Randy Truong"
date: "2022-12-29"
output: html_document
---

# Installing Packages 

```{r loadPackages, message=FALSE}
library(tidyverse)
library(readr)
```

# Data Setup  

## TODOS: 
-> Import all of the CSVs - **DONE**   
-> Reduce the DFs to essential rows - **DONE**   
  - CVD Diagnosis    
  - Diabetes diagnosis    
  - log(hsCRP)   
  - Abnormally high CRP rates    
  - # of Observations      
        
-> Create a race column - **DONE**    
-> Join all of the datasets together - **DONE**   
-> Create Plots     

## Important Terminology: 
CVD -> Cardiovascular Disease           
CRP -> C-Reactive Protein           

```{r loadData, message=FALSE}
fAsianDF <- read_csv("./Data/fData/F_Asian.csv")
fHispanicDF <- read_csv("./Data/fData/F_hispanic.csv")
fBlackDF <- read_csv("./Data/fData/F_non-hispanic_black.csv")
fWhiteDF <- read_csv("./Data/fData/F_non-hispanic_white.csv")

```

# Filtering of Data Sets 

In the datasets, we mostly just care about the rows "Diagnosed with CVD" (1),  
"Ever diagnosed with diabetes W5" (11), "log10(hsCRP) W4" (24), "High hsCRP (>3 mg/L) W4" (25), and "Observations" (34).

```{r filterMaleAsian, message=FALSE}

# Remember to rename first row 
fAsianDFClean <- mutate(fAsianDF, Race = "Asian") %>% 
  slice(1, 11, 23, 24, 25, 34) 

fHispanicDFClean <- mutate(fHispanicDF, Race = "Hispanic") %>% 
  slice(1, 11, 23, 24, 25, 34) 

fBlackDFClean <- mutate(fBlackDF, Race = "Black") %>% 
  slice(1, 11, 23, 24, 25, 34) 

fWhiteDFClean <- mutate(fWhiteDF, Race = "White") %>% 
  slice(1, 11, 23, 24, 25, 34) 

```

```{r createGiantFrame, message=FALSE}
raceList1 <- rbind(fAsianDFClean, fHispanicDFClean)
raceList2 <- rbind(fBlackDFClean, fWhiteDFClean)

finalRaceListF <- rbind(raceList1, raceList2)

```


# Data Visualization  

## TODOS: 

-> Visualize hsCRP rate by race     
-> Visualize the frequency of CVD and the frequency of diabetes diagnoses    
-> Compare log(hsCRP) to CVD / Compare log(hsCRP) to Diabetes    

```{r crpByRace}
# Note: The matrix must be race x summary statistics in order of (min, Q1, Q2, Q3, max)
#         [min] [q1] [q2] [q3] [max]
# [Asian]
# [Hispanic]
# [Black]
# [White]

finalRaceListF_CRP <- finalRaceListF %>% 
  slice(3, 9, 15, 21) 

fStats <- list(stats=t(cbind(finalRaceListF_CRP$Minimum,
                 finalRaceListF_CRP$`Percentile 25`,
                 finalRaceListF_CRP$Median,
                 finalRaceListF_CRP$`Percentile 75`,
                 finalRaceListF_CRP$Maximum)),
               n=c(finalRaceListF_CRP$Count[1],
                   finalRaceListF_CRP$Count[2],
                   finalRaceListF_CRP$Count[3],
                   finalRaceListF_CRP$Count[4]))
                

bxp(fStats, show.names = FALSE, main="Measure of hsCRP by Race (female)", xlab="hsCRP(mg/L)", ylab="Race", ylim = c(0,10), horizontal = TRUE)
axis(2, at = 1:4, labels = c("Asian", "Hispanic", "Black", "White"))

```
```{r diseaseFrequency}
# Frequency of Cardiovascular Disease (CVD) diagnoses 
# Bar graph 

```

```{r crpByDisease,echo=FALSE}
# Relationship between hsCRP and Cardiovascular Disease
# Bar graph 

# Relationship between hsCRP and Diabetes
# Bar graph 

```
